An efficient audio coding scheme for large scale acoustic monitoring

### 2017-06-12 paper/article.tex
second ml pass

\newcommand{\ml}[1]{\textcolor{blue}{ Mathieu: #1}}

@@

The present work is constrained to allow both acoustic monitoring and sound event recognition in urban soundscapes. Here we briefly review state-of-the-art methods in both subjects to guide our choices.

@@

The considered sensor network primarily aims at monitoring urban soundscapes, that is, continously assessing their content and impact on the population. This is typically enabled by measuring energetic acoustic indicators such as the equivalent sound pressure level $L_{eq}$ in dB SPL or its A-weighted equivalent $L_{Aeq}$ in dBA. However, while these indicators have proven to correlate well with perceptual evaluations for negative impact sound environments \cite{gozalo2015}, they fail to fully describe soundscapes \cite{rychtarikova2013}. Many other variables can be derived to better account for previously implicit properties \cite{can2016} including percentile values or time evolution. Studies have been conducted to select relevant subsets of descriptors in sound environment characterization \cite{can2015, brocolini2013, nilsson2007}.\\

@@

Another major interest of this work is sound event recognition. This problem has been the subject of extensive research in the past on speech \cite{anusuya2009}, music \cite{tzanetakis2002}, and lately more complex scenes in which the current work falls. Studied classification methods are diverse, ranging from time-dependant modeling with HMMs \cite{ntalampiras2014} to "bag-of-frames" approaches \cite{aucouturier2007, foggia2015}. Common architectures include learning-based classifiers such as support vector machines (SVM) \cite{kumar2016}, gaussian mixture models (GMM) \cite{radhakrishnan2005} or neural networks \cite{salamon2017, piczak2015}. However, the selection of relevant features is still an open debate. The most used are certainly spectral \cite{khunarsal2013} or cepstral \cite{couvreur2004} representations of the signal. Among them, mel spectrograms and their cepstral-domain derivation, the mel-frequency cepstrum coefficients (MFCC), are the most recurrent. These representations effectively try to model the human cochlear response to sounds by grouping frequency components around critical bands in a logarithmic scale, thus bearing important physical significance. They may also be exploited together with features computed in other domains to better model signal properties. For instance, \cite{cai2006} adds spectral features related to harmonicity and salient frequencies, and \cite{chu2009} uses a matching pursuit (MP) algorithm to deduce time-domaid features. Another promising solution is feature engineering via unsupervised learning, which \cite{salamon2015-2} implements with a k-means clustering technique. Alternative data representations such as the scattering transform \cite{bauge2013} show good results in environmental sound classification tasks \cite{salamon2015}.\\
A more detailed review of used methods is available in \cite{chachada2013}.

@@

The technical constraints of size inherent to the studied setup make it impossible to transmit raw audio recordings. Following the considerations briefly exposed in the previous section, we conclude that third-octave bands are a reasonable choice. As mentionned, they provide advantages in both acoustic monitoring and data volume per measurement period. The physical content is also close to that of mel spectrograms, being just another filterbank transform with logarithmic frequency scaling.\\

@@

The first step is to represent the signal in the frequency domain using a short-term Fourier transform (STFT). The continuously sampled audio is segmented into 125~ms frames to provide "fast" acoustic measurements. The data is zero-passed to the next power of two to allow best fft performances. No overlap is used as it doubles both computational and memory costs and is not required for our application. A rectangular window is applied to ensure energy conservation in a given analysis frame. The phase is unimportant to third-octave analysis, thus it is discarded. Similarly, negative frequency components contain the same information as positive ones because the base signal is real. We then compute third-octave bands from the squared spectral magnitude by matrix multiplication with the frequency weights proposed in \cite{antoni2010} for $l = 2$\footnote{When implementing said method, we found that squaring the $\phi_l(p), l \geq 1$ intermediate function (p. 887) was the correct approach to meet cutoff frequencies requirements, and we believe this was the author's original intention.}. We include analysis for bands $i$ from -17 to 13, ie. center frequencies $f_i$ ranging from $20~Hz$ to $20~kHz$ with $f_0 = 1~kHz$. This means the main representation is composed of 31 values sampled every 125~ms.\\

Alternatively, we also implement the mel filterbank for testing purposes. Mel spectrograms can be used as a baseline method to evaluate the relative performance of third-octave bands in sound event recognition. Furthermore, this representation is not bound by constraints such as conservation of energy, specific framing parameters or fixed resolution. It offers a more flexible tool to study intelligibility alteration and coder performance, both of which we take interest in. Mel spectrograms are computed with the \textit{rastamat} library \cite{ellis2005}. STFT analysis frames are obtained by applying a Hann window on 23.2~ms of signal with 50\% (11.6~ms) overlap to allow for efficient phase recovery in further tests.

@@

A Huffman coding scheme \cite{huffman1952} is then used to further reduce data dimensionality. As in most entropy coding algorithms, the efficiency of this technique depends on two important factors. A reduced amount of symbols in the dictionnary yields smaller code size on lower probability symbols. Lower data entropy, ie. lower average information content of the signal distribution, also increases performance. The first is generally obtained by applying a quantization process to the signal, whereas the second is directly linked with the probability density function (PDF) of these symbols. It is defined as $H = -\sum\limits_i p_ilog_n(p_i)$, where $p_i$ is the probability of appearance of a given symbol and $n$ the numerical base in which information is represented. As such, entropy decreases when very few symbols have a very high probability of appearance and is maximum for a uniform distribution. Considering an estimated PDF for our data as shown in Figure~\ref{fig:pdf}a, immediately applying a linear quantization clearly results in most of the information being lost. We therefore spread the PDF by taking the logarithm of the representation. A linear quantization is then applied with $2^{q-1}-1$ output values to obtain Figure~\ref{fig:pdf}b. Finally, we use a $\Delta$-encoding algorithm along the time dimension to reduce redundancies. This also effectively concentrates higher probabilities on symbols around zero (Figure~\ref{fig:pdf}c). This yields a higher amount of symbols at $2^q-1$ but a nevertheless smaller entropy. In the example shown here, the former is $H_{log} = 6.24~sh$ and the latter $H_{\Delta} = 3.54~sh$. Huffman encoding is then computed with either a frame-specific symbol-code dictionnary or a constant one generated from an entire dataset. A comparison of both methods is shown in Figure~\ref{fig:dict_comp}. For most short texture frame durations, we found the local Huffman algorithm to be much faster. In fact, encoding complexity is function of the amount of dictionnary elements. If it is specific to a given data packet, then it does not need to contain every possible value. This gain seem to outweight the additional complexity induced by the tree generation algorithm. Alternatively, short texture frames or single analysis windows can be encoded using one general dictionnary to improve bitrate. This choice depends on which of the size and complexity parameters must be reduced most.\\

@@

The decoding process is quite straightforward, as both Huffman and $\Delta$ compression are lossless and directly reversible.

The entire coder scheme is summarized in Figure~\ref{fig:scheme}.

@@

The coder's output bitrate as well as additional measurement error is computed for both data representations, it is expressed in dB and compared to the IEC 61672-1:2013 standard on sound level meter tolerance. This is to verify that the quantization step yields a maximum absolute error inferior to the precision of the recording device. To provide relevant statistics, these quantities are estimated for a total of 53 10-minutes texture frames.

@@

The UrbanSound8k dataset\cite{salamon2014} features about 9 hours of urban environmental sounds recordings separated in 8732 wave files ranging from 1 to 4 seconds each. The recordings are labeled in 10 classes (air conditioner, car horn, children playing, dog bark, drilling, enginge idling, gun shot, jackhammer, siren, street music), and distributed in 10 independant folds. A method and baseline results are also provided for the classification task. It is used in all the following experiments with the exception of intelligibility computation for which a small dataset of high-quality voice recordings is preferred. Audio files are resampled 44.1 samples per second, normalized and reduced to a single channel.\\

We first evaluate the global loss of information by implementing the four classification models proposed in \cite{salamon2014}: a support vector machine with a radial basis function kernel, a random forest classifier with 500 trees, a decision tree and a k-nearest neighbors classifier with $k = 5$. Values for the SVM parameter $C$ and RBF kernel variance $\sigma^2$ are found with a grid search. We apply a discrete cosine transform to the critical band signal representation to obtain cepstra (known as Mel Frequency Cepstrum Coefficients or MFCC in the first case) of which we conserve the 25 first coefficients, and summarize along time with the mean, variance, skewness, kurtosis, minimum, maximum, median, derivative mean and variance, second order derivative mean and variance. The feature vector is thus comprised of 275 values to reproduce available results and compare with our own. Models are trained for each setup using the 10-fold method, that is, every combination of testing one fold on models trained with the other nine.

@@

Intelligibility in decoded and reconstructed audio is also one of the main concerns of this study. We conduct preliminary perceptive tests to ensure that a clean speech recording is made unintelligible, implying the same results in usually noisier urban soundscapes. We also provide the computation of two objective metrics: the Coherence Speech Intelligibility Index\cite{kates2005} (CSII) and frequency-eighted segmental SNR\cite{hu2008} (fwSNRseg) by comparing original unaltered samples and recovered audio. These two indicators have been shown to correlate well with perceptual tests\cite{ma2009}, although results are only available for lesser degradations and distortions.\\

These metrics require a recovery of the time-domain signal. We must first recover a linearly-scaled spectrogram from the band-focused representations. This can be achieved by multiplying by the scaled transpose of the forward transformation matrix. A loss of resolution increasing with frequency is induced due to the shape of the transformation, as illustrated by Figure~\ref{fig:freq}. Signal phase is approximated by either white noise spectrogram scaling or a Griffin\& Lim algorithm \cite{griffin1984}, and the signal is retrieved using overlap-add. When using overlap while computing third-octave bands, one can also avoid framing effects produced by rectangular windowing by convoluting the signal with another windowing function prior to inverse-STFT computation.\\

>>>>

\newcommand{\ml}[1]{\textcolor{blue}{ Mathieu: #1}}
\newcommand{\fg}[1]{\textcolor{red}{ Felix: #1}}

@@

The present work is constrained to allow both acoustic monitoring and sound event recognition in urban soundscapes. Here we briefly review state-of-the-art methods in both subjects in order to better motivate our choices.

@@

The considered sensor network primarily aims at monitoring urban soundscapes, that is, continously assessing their content and impact on the population. In the literature, this is typically enabled by measuring energetic acoustic indicators such as the equivalent sound pressure level $L_{eq}$ in dB SPL or its A-weighted equivalent $L_{Aeq}$ in dBA. However, while these indicators have proven to correlate well with perceptual evaluations for negative impact sound environments \cite{gozalo2015}, they are not sufficient to fully describe soundscapes \cite{rychtarikova2013}. Many other variables can be derived to better account for previously implicit properties \cite{can2016} including percentile values or time evolution. Studies have been conducted to select relevant subsets of descriptors in sound environment characterization \cite{can2015, brocolini2013, nilsson2007}.\\

@@

Eventhough the above discussed indicators provide more richness than $L_{eq}$, recent studies show that abstract statistics are still not sufficient to fully model the human perception of soundscapes. The recognition of sources of interest is thus of importance. As far as the encoding scheme is concerned, it is thus important that the encoded data allows the computation of the above cited indicators but also the recognition of sources of interest with state of the art methods.

The recognition of sources of interest from audio streams has been the subject of extensive research in the past on speech \cite{anusuya2009}, music \cite{tzanetakis2002}, and lately more complex scenes in which the current work falls. Studied classification methods are diverse, ranging from time-dependant modeling with HMMs \cite{ntalampiras2014} to "bag-of-frames" approaches \cite{aucouturier2007, foggia2015}. Common architectures include learning-based classifiers such as support vector machines (SVM) \cite{kumar2016}, gaussian mixture models (GMM) \cite{radhakrishnan2005} or neural networks \cite{salamon2017, piczak2015}. However, the selection of relevant features is still an open debate. The most used are certainly spectral \cite{khunarsal2013} or cepstral \cite{couvreur2004} representations of the signal. Among them, mel spectrograms and their cepstral-domain derivation, the mel-frequency cepstrum coefficients (MFCC), are the most recurrent. These representations effectively try to model the human cochlear response to sounds by grouping frequency components around critical bands in a logarithmic scale, thus bearing important physical significance. They may also be exploited together with features computed in other domains to better model signal properties. For instance, \cite{cai2006} adds spectral features related to harmonicity and salient frequencies, and \cite{chu2009} uses a matching pursuit (MP) algorithm to deduce time-domaid features. Another promising solution is feature engineering via unsupervised learning, which \cite{salamon2015-2} implements with a k-means clustering technique. Alternative data representations such as the scattering transform \cite{bauge2013} show good results in environmental sound classification tasks \cite{salamon2015}.\\
A more detailed review of used methods is available in \cite{chachada2013}.

@@

The technical constraints of size inherent to the studied setup make it impossible to transmit raw audio recordings. Following the considerations briefly exposed in the previous section, we conclude that third-octave bands are a reasonable choice. As mentionned, they provide advantages in both acoustic monitoring quality and reduced data volume per measurement period. The physical content is also close to that of mel spectrograms, being just another filterbank transform with logarithmic frequency scaling.\\

@@

The first step is to represent the signal in the frequency domain using a short-term Fourier transform (STFT). The continuously sampled audio is segmented into 125~ms frames to provide "fast" acoustic measurements. The data is zero-padded to the next power of two to allow best fft performances. No overlap is used as it doubles both computational and memory costs and is not required for our application. A rectangular window is thus applied to ensure energy conservation in a given analysis frame. The phase is unimportant to third-octave analysis, thus it is discarded. Similarly, negative frequency components contain the same information as positive ones because the base signal is real. We then compute third-octave bands from the squared spectral magnitude by matrix multiplication with the frequency weights proposed in \cite{antoni2010} for $l = 2$\footnote{When implementing said method, we found that squaring the $\phi_l(p), l \geq 1$ intermediate function (p. 887) was the correct approach to meet cutoff frequencies requirements, and we believe this was the author's original intention.}. We include analysis for bands $i$ from -17 to 13, ie. center frequencies $f_i$ ranging from $20~Hz$ to $20~kHz$ with $f_0 = 1~kHz$. This means the main representation is composed of 31 values sampled every 125~ms.\\

Alternatively, we also implement the mel filterbank for comparison purposes, as the mel filterbank is the root representation of many state of the art classification algorithms. Mel spectrograms can thus be used as a baseline method to evaluate the relative performance of third-octave bands in sound event recognition.

As the Mel representation is not bounded by constraints such as conservation of energy, specific framing parameters or fixed resolution, it also offers a more flexible tool to study intelligibility alteration and coder performance, both of which we take interest in.

Mel spectrograms are computed with the \textit{rastamat} library \cite{ellis2005}. STFT analysis frames are obtained by applying a Hann window on 23.2~ms of signal with 50\% (11.6~ms) overlap to allow for efficient phase recovery in further tests.

@@

A Huffman coding scheme \cite{huffman1952} is then used to further reduce data dimensionality. As in most entropy coding algorithms, the efficiency of this technique depends on two important factors. A reduced amount of symbols in the dictionnary yields smaller code size on lower probability symbols. Lower data entropy, ie. lower average information content of the signal distribution, also increases performance. The first is generally obtained by applying a quantization process to the signal, whereas the second is directly linked with the probability density function (PDF) of these symbols. It is defined as $H = -\sum\limits_i p_ilog_n(p_i)$, where $p_i$ is the probability of appearance of a given symbol and $n$ the numerical base in which information is represented.

As such, entropy decreases when very few symbols have a very high probability of appearance and is maximum for a uniform distribution. Considering an estimated PDF for our data as shown in Figure~\ref{fig:pdf}a, immediately applying a linear quantization clearly results in most of the information being lost. We therefore spread the PDF by taking the logarithm of the representation. A linear quantization is then applied with $2^{q-1}-1$ output values to obtain Figure~\ref{fig:pdf}b. Finally, we use a $\Delta$-encoding algorithm along the time dimension to reduce redundancies. This also effectively concentrates higher probabilities on symbols around zero (Figure~\ref{fig:pdf}c). This yields a higher amount of symbols at $2^q-1$ but a nevertheless smaller entropy.

In the example shown here, the former is $H_{log} = 6.24~sh$ and the latter $H_{\Delta} = 3.54~sh$. Huffman encoding is then computed with either a frame-specific symbol-code dictionnary or a constant one generated from an entire dataset. A comparison of both methods is shown in Figure~\ref{fig:dict_comp}. For most short texture frame durations, we found the local Huffman algorithm to be much faster. In fact, encoding complexity is function of the amount of dictionnary elements. If it is specific to a given data packet, then it does not need to contain every possible value. This gain seem to outweight the additional complexity induced by the tree generation algorithm. Alternatively, short texture frames or single analysis windows can be encoded using one general dictionnary to improve bitrate. This choice depends on which of the size and complexity parameters must be reduced most.\\

@@

The decoding process is quite straightforward, as both Huffman and $\Delta$ compression are lossless and directly reversible. The entire coder scheme is summarized in Figure~\ref{fig:scheme}.

@@

The coder's output bitrate as well as additional measurement error is computed for both data representations : third-octave and Mel. It is expressed in dB and compared to the IEC 61672-1:2013 standard on sound level meter tolerance. This is to verify that the quantization step yields a maximum absolute error inferior to the precision of the target recording device. To provide relevant statistics, these quantities are estimated for a total of 53 10-minutes texture frames \ml{describe dataset, etc}.

@@

In order to ensure that the proposed scheme allows the recognition of events using state of the art methods, we consider the UrbanSound8k dataset\cite{salamon2014}. It features about 9 hours of urban environmental sounds recordings separated in 8732 wave files ranging from 1 to 4 seconds each. The recordings are labeled in 10 classes (air conditioner, car horn, children playing, dog bark, drilling, enginge idling, gun shot, jackhammer, siren, street music), and distributed in 10 independant folds. A method and baseline results are also provided for the classification task. It is used in all the following experiments with the exception of intelligibility computation for which a small dataset of high-quality voice recordings is preferred. Audio files are resampled 44.1 samples per second, normalized and reduced to a single channel.\\

We first evaluate the global loss of information by implementing the four classification models proposed in \cite{salamon2014}: a support vector machine with a radial basis function kernel, a random forest classifier with 500 trees, a decision tree and a k-nearest neighbors classifier with $k = 5$. Values for the SVM parameter $C$ and RBF kernel variance $\sigma^2$ are found with a grid search. We apply a discrete cosine transform to the critical band signal representation to obtain cepstra (known as Mel Frequency Cepstrum Coefficients or MFCC in the first case) of which we conserve the 25 first coefficients, and summarize along time with the mean, variance, skewness, kurtosis, minimum, maximum, median, derivative mean and variance, second order derivative mean and variance. The feature vector is thus comprised of 275 values to reproduce available results and compare with our own. Models are trained for each setup using a 10-fold cross validation method \ml{provided by the authors of \cite{salamon2014} ?}, that is, every combination of testing one fold on models trained with the other nine.

@@

Intelligibility in decoded and reconstructed audio is also oan concerns of this study. It is indeed important that the produced scheme maintains a high level of recognition of acoustic events but, as the dat is transmitted over the network and potentially stored, the level of intelligibility of the decoded stream should be as low as possible.

%conduct thus preliminary perceptive tests to ensure that a clean speech recording is made unintelligible, implying the same results in usually noisier urban soundscapes. We also

We thus provide the computation of two objective metrics: the Coherence Speech Intelligibility Index\cite{kates2005} (CSII) and frequency-weighted segmental SNR\cite{hu2008} (fwSNRseg) by comparing original unaltered samples and recovered audio. These two indicators have been shown to correlate well with perceptual tests\cite{ma2009}, although results are only available for lesser degradations and distortions than the ones we consider in this study.\\

These metrics require a recovery of the time-domain signal. To do so, a linearly-scaled spectrogram from the band-focused representations is estimated. This estimation can be achieved by multiplying by the scaled transpose of the forward transformation matrix. A loss of resolution increasing with frequency is induced due to the shape of the transformation, as illustrated by Figure~\ref{fig:freq}. Signal phase is approximated by either white noise spectrogram scaling or a Griffin\& Lim algorithm \cite{griffin1984}, and the signal is retrieved using concatenation or overlap-add. When using overlap to compute third-octave bands, one can also avoid framing effects produced by rectangular windowing by convoluting the signal with another windowing function prior to inverse-STFT computation.\\
